---
title: "Logistic regression assignment, group 3"  
author: Elizabeth Machkovska, Letícia Marçal Russo, Madio Seck, William Schaafsma 
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
       highlight: textmate
       theme: flatly
       number_sections: yes
       toc: yes
       toc_float:
         collapsed: yes
         smooth_scroll: no
---

# Context 


# Packages
```{r libraries, warning = FALSE, message = FALSE}
library(ggplot2)

```

# Research question and relevance 


# Dataset and variables

# Inspecting the dataset and cleaning up

## Missing values 
## Looking for outliers

# Descriptive analysis

# Correlation 

  
# Set seed
Setting seed to generate a reproducible result.
```{r }
set.seed(42)

```

# Best model

# Comparing models

# Cross Validation

# Confusion Matrix

# Visualization of the model 

# Checking assumptions 

## Residuals vs Fitted model

## Normal Q-Q plot

## Scale-Location 

## Cook's distance

## Residuals vs Leverage

# Interpreting the final model

# Answering the research question 

# Discussion & Limitations


# Sources

Hastie, T., James, G., Tibshirani, R. & Witten, D. (2021). *An Introduction to Statistical Learning with Applications in R*. (2nd edition). Springer. Retrieved from https://hastie.su.domains/ISLR2/ISLRv2_website.pdf

Haira, K. (2018, 17 May). *dplyr - advanced-joining* Retrieved from https://rpubs.com/KunalHaria/390063.

Kassambara, A. (2018). Linear Regression Assumptions and Diagnostics in R: Essentials. In: *Statistical Tools for High-Throughput Data Analysis.* (1st edition). Retrieved from http://www.sthda.com/english/articles/39-regression-model-diagnostics/161-linear-regression-assumptions-and-diagnostics-in-r-essentials/#diagnostic-plots.

Klein, A. G., Gerhard, C., Büchner, R. D., Diestel, S. & Schermelleh-Engel, K. (2016). The detection of heteroscedasticity in regression models for psychological data. *Psychological Test and Assessment Modeling, 58*(4), 542-568.

Rousselet, G. A., & Wilcox, R. R. (2020). Reaction times and other skewed distributions: problems with the mean and the median. *Meta-Psychology, 4*. 
https://doi.org/10.15626/MP.2019.1630




